Upwelling variability across the Benguela upwelling systems

### 2019-09-23 Write_up/Paper1_writeup.Rmd
merge

Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa. Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions. Specific wind characteristics were measured, namely; wind direction (dirw) and wind speed (spw). Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data. The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters. ERA5 hourly data on pressure levels was used as a second wind source. This dataset comprised of u and v variables from 1973 to present.

With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested. To do this, it was important to identify upwelling.

@@

In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The coastal angle is the angle of each of the sites perpendicular to the coastline. The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur. Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value was ment to be. Here, it was specified that the minimum duration for upwelling was 1 day with a threshold value obtained from the upwelling index as 1. In order to obtain a more approved threshold value, a percentile value was determined. To determine changes within the upwelling values, the 'detect event' function within the heatwaveR package was used. Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile. Values obtained from the upwelling index were used as a second threshold. An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data. Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months. For the use of the heatwaveR package it was best to use a time series of 30years.

>>>>

Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling, consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa. Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions. Specific wind characteristics were measured, namely; wind direction (dirw) and wind speed (spw). Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data. The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters.

With temperature and wind values now corresponding to their respective sites and dates, the hypothesis examining upwelling patterns at different distances from the coastline could now be tested. To do this, it was first important to identify upwelling.

@@

In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The coastal angle is the angle of each of the sites perpendicular to the coastline. The index relies heavily on wind speed and direction data in order to determine the intensity of upwelling likely to occur. Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value was ment to be. Here, it was specified that the minimum duration for upwelling was 1 day with a threshold value obtained from the upwelling index as 1. In order to obtain a more approved threshold value, a percentile value was determined. To determine changes within the upwelling values, the 'detect event' function within the heatwaveR package was used. Thereafter, to determine the upwelling event metrics (frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile. Values obtained from the upwelling index were used as a second threshold. An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data. Because upwelling events were calculated relative to percentile exceedances, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occurred any time of the year with upwelling shown to be more dominant during spring and summer months. For the application of the heatwaveR package it was best to make use a time series of 30 years.
